---
title: "Replication of Bond et al 2001"
author: "Claudius"
date: "5/27/2021"
output:
  pdf_document:
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(plm)
library(here)
library(pder)
library(data.table)
library(stargazer)
```

```{r}
data("DemocracyIncome", package = "pder")
```

```{r, eval=FALSE, echo=FALSE}
DemocracyIncome2 <- fread(
  here("data/acemoglu2001.csv"), 
  sep = ";", dec = ",", 
  select = c(
    "sample"="double", 
    "country"="character", 
    "year"="factor", 
    "year_numeric"="double", 
    "fhpolrigaug"="double", 
    "lrgdpch"="double")
  ) %>%
  rename(
    income=fhpolrigaug,
    democracy=lrgdpch
  ) %>%
  mutate(
    year2=as.double(year)
  )
```

## General setting

The econometric model is given as follows:

$$d_{it} = \alpha d_{it-1} + \gamma y_{it-1} + \boldsymbol{x}'\boldsymbol{\beta} + \mu_t + \delta_i + u_{it}$$

where $d_{it}$ represents the democracy index of country $i$ in year $t$,
$\alpha$ is the autoregressive parameter to be estimated (and which accounts 
for persistency and mean-reverting dynamics).
$\gamma$ is the parameter on the impact of lagged income $y$.
It is the main parameter of interest in this study.
$\boldsymbol{x}'$ contains additional controls,
$\mu_t$ are time, and $\delta_i$ are country fixed effects.
The idiosyncratic error is $u_{it}$.

In the following we replicate the first five models of table 2 in the main paper.

## Replication of the main results

### Pooled OLS

The POLS model does not consider any country fixed effects (i.e. excludes 
$\delta_i$):

```{r}
pols_model <- plm(
  democracy ~ lag(democracy) + lag(income), model = "within", 
  data = DemocracyIncome, #index = c("country", "year_numeric"),
  subset = sample==1, effect = "time")
pols_model
```

```{r, eval=FALSE, echo=FALSE}
pols_model <- plm(
  democracy ~ lag(democracy) + lag(income), model = "within", 
  data = DemocracyIncome2, index = c("country", "year2"),
  subset = sample==1, effect = "time")
pols_model
```

### Within model

The *within*-model is the same as the POLS model but with $\delta_i$
included:

```{r}
within_model <- plm(
  democracy ~ lag(democracy) + lag(income), model = "within", 
  data = DemocracyIncome, subset = sample==1, effect = "twoways")
within_model
```

## Anderson-Hsiao

This is an estimation estimates the model from above in differences:

$$\Delta d_{ut} = \alpha \Delta d_{it-1} + \gamma\Delta y_{it-1} + \Delta + \boldsymbol{x}'\boldsymbol{\beta} + \Delta  \mu_t + \Delta u_{it}$$

Using differences eliminates the country specific error component $\delta_i$.
Nevertheless, $\Delta d_{it-1}$ and $\Delta y_{it-1}$ are correlated with 
$u_{it}$ by construction.
To remedy this, $y_{it-1}$ is used as an instrument for $\Delta y_{it-1}$ and 
$d_{it-2}$ is used as an instrument for $\Delta d_{it-1}$:

```{r}
ahsiao_model <- plm(
  diff(democracy) ~ lag(diff(democracy)) + lag(diff(income)) | 
    lag(democracy, 2) + lag(income, 2), 
  data = DemocracyIncome, subset = sample==1, 
  model = "within", effect = "time")
ahsiao_model
```

### Arellano Bond

```{r}
abond_model <- pgmm(
  democracy ~ lag(democracy) + lag(income) | lag(democracy, 2:99) | lag(income, 2), 
  data = DemocracyIncome, subset = sample==1, effect = "twoways", model = "twostep")
summary(abond_model)
```


```{r, results='asis'}
stargazer(
  pols_model, within_model, ahsiao_model, abond_model)
```

